---
title: "Responses Analysis"
author: "Ryan Wesslen"
date: "July 20, 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Load & Pre-processing

### Read in the File

```{r}
#install.packages("tidyverse")
library(tidyverse)

responses <- read_csv("./data/userForms.csv")

table(responses$fairnessRadios)

# rename columns -- should be a better way
names <- colnames(responses) 
names[14:18] <- c("accountsImportance", 
                  "networkImportance", 
                  "tweetsImportance",
                  "geoImportance", 
                  "entityImportance")
colnames(responses) <- names
```

Let's see how many responses by user.

```{r}
table(responses$user_id)

responses$user_id <- toupper(responses$user_id)
```

Let's now remove the test accounts.

```{r}
excludeRows <- grep("TEST", responses$user_id)
responses <- responses[-excludeRows,]
```

Also notice that some users have > 8 responses. These are cases when they responded twice.

Let's identify then drop them.

```{r}
# order by user, account, then time string (note descending so 1 = last, 2 = 2nd last, ...)
responses <- responses[order(responses$user_id, 
                             responses$account_id, 
                             desc(responses$currentTimeString)),]

# create user-account key
responses$responseKey <- paste0(responses$user_id,"+",responses$account_id)

# create count
responses$order <- sequence(rle(responses$responseKey)$lengths)

# remove duplicates
responses <- filter(responses, order == 1)
```

FYI CC-17 only reported for 7.

### Reformat Columns

Create a new column that says the truth based on a mapping table.

```{r}
vlookup <- function(this, data, key, value){
  m <- match(this, data[[key]])
  data[[value]][m]
}

map <- tibble(account_id = c("04","39","59","83","02","48","56","71"), 
              truth = c("fake","fake","fake","fake","real","real","real","real"),
              accountName = c("@AddInfoOrg","@NatCounterPunch","@SGTreport","@zerohedge",
                                  "@ABC","@nytimes","@Reuters","@TIME"))

responses <- responses %>%
              mutate(truth = vlookup(account_id, map, "account_id", "truth"),
                     accountName = vlookup(account_id, map, "account_id", "accountName" ),
                     group = case_when(
                       substr(user_id,1,2)=="CC" ~ "Control",
                       substr(user_id,1,2)=="TC" ~ "Confirm",
                       TRUE ~ "Disconfirm"
                     ),
                     accuracy = ifelse(verifyRadios==truth,1L,0L),
                     fake = ifelse(verifyRadios=="fake",1L,0L))
```

### Regression

```{r}
accuracy <- mean(as.integer(responses$accuracy))
responses$accuracy <- as.factor(responses$accuracy)

recoded <- function(var){
  case_when(
    var == -1 ~ 0,
    var == 0 ~ -1,
    var == 1 ~ 1
  )
}

# recoding responses so -1 = No, 0 = n/a, and 1 = Yes
responses$socialRadios <- recoded(responses$socialRadios)
responses$tweetRateRadios <- recoded(responses$tweetRateRadios)
responses$fairnessRadios <- recoded(responses$fairnessRadios)
responses$entityRadios <- recoded(responses$entityRadios)
responses$fearRadios <- recoded(responses$fearRadios)
responses$opiniatedRadios <- recoded(responses$opiniatedRadios)

# social network, bias/fairness, fear/negative => point to real
# tweet rate, entity, opininated => point to fake

realCue <- function(var, actual){
  case_when(
    var == 1 & actual == "real" ~ 1L,
    var == 1 & actual == "fake" ~ -1L,
    var == -1 & actual == "real" ~ -1L,
    var == -1 & actual == "fake" ~ 1L,
    var == 0 ~ 0L
  )
}

fakeCue <- function(var, actual){
  case_when(
    var == 1 & actual == "real" ~ 1L,
    var == 1 & actual == "fake" ~ 1L,
    var == -1 & actual == "real" ~ 1L,
    var == -1 & actual == "fake" ~ -1L,
    var == 0 ~ 0L
  )
}

responses <- responses %>%
  mutate(socialRadios2 = realCue(socialRadios, truth),
         tweetRateRadios2 = fakeCue(tweetRateRadios, truth),
         fairnessRadios2 = realCue(fairnessRadios, truth),
         entityRadios2 = fakeCue(entityRadios, truth),
         fearRadios2 = realCue(fearRadios, truth),
         opiniatedRadios2 = fakeCue(opiniatedRadios, truth))
```

### Accuracy Dependent Variable

```{r}
form <- "accuracy ~ socialRadios2 + tweetRateRadios2 + fairnessRadios2 + entityRadios2 + fearRadios2 + opiniatedRadios2"

fit <- glm(form, 
           data = responses, 
           family = "binomial")

summary(fit)

exp(cbind(OR = coef(fit), confint(fit)))

pred1 <- predict(fit, type= "response") > accuracy

table(pred1, responses$accuracy)
```

```{r}
# form 2: cue ratings + account name
responses$accountName <- as.factor(responses$accountName)
form2 <- paste(form, "+ accountName")

fit2 <- glm(form2, 
           data = responses, 
           family = "binomial")

summary(fit2)

exp(cbind(OR = coef(fit2), confint(fit2)))

pred2 <- predict(fit2, type= "response") > accuracy 

table(pred2, responses$accuracy)
```

#### form 3: cue ratings + account name + confidence

```{r}
form3 <- paste(form2,"+ confidence")

fit3 <- glm(form3,
            data = responses,
            family = "binomial")

summary(fit3)

exp(cbind(OR = coef(fit3), confint(fit3)))

pred3 <- predict(fit3, type= "response") > accuracy 

table(pred3, responses$accuracy)
```

#### Form 4: Treatment Groups

```{r}
responses$group <- factor(responses$group, levels = c("Control","Confirm","Disconfirm"))

form4 <- paste(form3, "+ group")

fit4 <- glm(form4,
            data = responses,
            family = "binomial")

summary(fit4)

exp(cbind(OR = coef(fit4), confint(fit4)))

pred4 <- predict(fit4, type= "response") > accuracy 

table(pred4, responses$accuracy)
```

### Fake Dependent Variable

```{r}
# form 5: cue ratings
responses$fake <- as.factor(responses$fake)

form5 <- "fake ~ socialRadios2 + tweetRateRadios2 + fairnessRadios2 + entityRadios2 + fearRadios2 + opiniatedRadios2"

fit5 <- glm(form5, 
           data = responses, 
           family = "binomial")

summary(fit5)

exp(cbind(OR = coef(fit5), confint(fit5)))

pred5 <- predict(fit5, type= "response") > accuracy

table(pred5, responses$fake)

# form 2: cue ratings + account name
responses$accountName <- as.factor(responses$accountName)
form6 <- paste(form5, "+ accountName")

fit6 <- glm(form6, 
            data = responses, 
            family = "binomial")

summary(fit6)

exp(cbind(OR = coef(fit6), confint(fit6)))

pred6 <- predict(fit6, type= "response") > accuracy 

table(pred6, responses$accuracy)

# form 3: cue ratings + account name + confidence

form7 <- paste(form6,"+ confidence")

fit7 <- glm(form7,
            data = responses,
            family = "binomial")

summary(fit7)

exp(cbind(OR = coef(fit7), confint(fit7)))

pred7 <- predict(fit7, type= "response") > accuracy 

table(pred7, responses$fake)

# form 4: 

responses$group <- factor(responses$group, levels = c("Control","Confirm","Disconfirm"))

form8 <- paste(form7, "+ group")

fit8 <- glm(form8,
            data = responses,
            family = "binomial")

summary(fit8)

exp(cbind(OR = coef(fit8), confint(fit8)))

pred8 <- predict(fit8, type= "response") > accuracy 

table(pred8, responses$fake)
```

### Export output to CSV


```{r}
df1 <- as.tibble(exp(cbind(OR = coef(fit4), confint(fit4))))
df2 <- as.tibble(exp(cbind(OR = coef(fit8), confint(fit8))))

dfA <- tibble(Name = rownames(exp(cbind(OR = coef(fit4), confint(fit4)))),
              Accuracy = df1$OR,
              Veracity = df2$OR,
              AccL = df1$`2.5 %`,
              AccU = df1$`97.5 %`,
              VerL = df2$`2.5 %`,
              VerU = df2$`97.5 %`)

dfA <- dfA %>% mutate_each(funs(round(.,2)), -Name) 

write_csv(dfA, "./data/dfA.csv")

df3 <- as.tibble(exp(cbind(OR = coef(fit), confint(fit))))
df4 <- as.tibble(exp(cbind(OR = coef(fit5), confint(fit5))))

# big

dfB <- tibble(Name = rownames(exp(cbind(OR = coef(fit), confint(fit)))),
              Accuracy = df3$OR,
              Veracity = df4$OR,
              AccL = df3$`2.5 %`,
              AccU = df3$`97.5 %`,
              VerL = df4$`2.5 %`,
              VerU = df4$`97.5 %`)

dfB <- dfB %>% mutate_each(funs(round(.,2)), -Name) 

write_csv(dfB, "./data/dfB.csv")
```

## Session Info

```{r}
sessionInfo()
```

